Reanalysis Step 4: Main Effects Outcome Analysis

### 2025-08-28 reanalysis-1-data_preparation.qmd
fix ps and implement quarto rstudio friendly logging

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
setup_analysis(seed = 2025)
```

@@

```{r load-data}
df_combined <- readr::read_csv(reanalysis_data_file_path)
```

@@

```{r summary}
# Final summary
tibble(
  Metric = c(
    "Total observations",
    "Complete cases (no missing in key variables)", 
    "Missing outcome VA (to be imputed)",
    "Missing CCI or DCSI (to be imputed)",
    "Missing baseline VA (to be imputed)",
    "Patients who lapsed",
    "Patients with vision impairment (observed)"
  ),
  N = c(
    nrow(df_for_imputation),
    nrow(df_complete_original),
    sum(is.na(df_for_imputation$outcome_VA_logMAR)),
    sum(is.na(df_for_imputation$CCI) | is.na(df_for_imputation$DCSI)),
    sum(is.na(df_for_imputation$baseline_VA_logMAR)),
    sum(df_for_imputation$ever_lapse_binary == 1, na.rm = TRUE),
    sum(df_for_imputation$outcome_va_vi_binary == 1, na.rm = TRUE)
  )
) |>
  kable(caption = "Dataset summary for reanalysis with imputation")
```

>>>>

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
setup_analysis(seed = 2025)

# Initialize logging
log_file <- init_log("data_preparation")
```

@@

```{r load-data}
log_message(paste0("Loading data from: ", reanalysis_data_file_path))
df_combined <- readr::read_csv(reanalysis_data_file_path)
log_message(paste0("Loaded ", nrow(df_combined), " observations"))
```

@@

```{r summary}
# Final summary
tibble(
  Metric = c(
    "Total observations",
    "Complete cases (no missing in key variables)", 
    "Missing outcome VA (to be imputed)",
    "Missing CCI or DCSI (to be imputed)",
    "Missing baseline VA (to be imputed)",
    "Patients who lapsed",
    "Patients with vision impairment (observed)"
  ),
  N = c(
    nrow(df_for_imputation),
    nrow(df_complete_original),
    sum(is.na(df_for_imputation$outcome_VA_logMAR)),
    sum(is.na(df_for_imputation$CCI) | is.na(df_for_imputation$DCSI)),
    sum(is.na(df_for_imputation$baseline_VA_logMAR)),
    sum(df_for_imputation$ever_lapse_binary == 1, na.rm = TRUE),
    sum(df_for_imputation$outcome_va_vi_binary == 1, na.rm = TRUE)
  )
) |>
  kable(caption = "Dataset summary for reanalysis with imputation")

# Finalize logging
log_message(paste0("Data preparation completed: ", nrow(df_for_imputation), " observations ready for imputation"))
final_log <- finalize_log(success = TRUE)
```

**Log file:** `r final_log`

### 2025-08-28 reanalysis-3-propensity_scores.qmd
fix ps and implement quarto rstudio friendly logging

```{r setup}
#| include: false
source("dependencies.R")
setup_analysis(seed = 2025)

@@

@@

# Load functions with formula definitions
source("functions.R")

# Get the ORIGINAL formula (does NOT include dr_severity)
ps_formula_original <- get_matching_formula()

@@

```{r ps-models}
# Initialize storage for PS results
twang_results <- list()

@@

for (i in 1:n_imputations) {
  # Run twang ps model - following analysis.Rmd approach
  twang_model <- ps(
    formula = ps_formula,
    data = data.frame(imputed_datasets[[i]]),  # Convert to data.frame as in analysis.Rmd
    n.trees = n_trees,
    interaction.depth = 3,
    shrinkage = 0.01,
    estimand = "ATT",
    stop.method = c("es.mean"),  # Using es.mean only for speed
    verbose = FALSE
  )

@@

  # Store results
  twang_results[[i]] <- list(
    model = twang_model,
    weights = weights,
    n_trees_used = twang_model$desc$es.mean$n.trees
  )

  # Track progress
  progress_df <- rbind(progress_df, data.frame(
    Imputation = i,
    Trees_Used = twang_model$desc$es.mean$n.trees,
    Weight_Min = round(min(weights), 3),
    Weight_Max = round(max(weights), 3)
  ))
}

@@

```{r save-results}
# Extract weights from all twang models for easy access in later steps
twang_weights <- list()
for (i in 1:n_imputations) {
  twang_weights[[i]] <- get.weights(twang_results[[i]], stop.method = "es.mean")
}

@@

**Propensity score results saved to:** `r file.path(reanalysis_data_dir, "ps_results_twang.rds")`
**Weights saved to:** `r file.path(reanalysis_data_dir, "twang_weights.rds")`

>>>>

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load logging functions and other utilities
setup_analysis(seed = 2025)

@@

# CACHING CONFIGURATION
use_cache <- TRUE  # Set to FALSE to force re-computation
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

@@

ps_formula_original <- get_matching_formula()

@@

```{r ps-models}
# Initialize logging
log_file <- init_log("propensity_scores")
log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Number of imputations to process: ", n_imputations))
log_message(paste0("Cache directory: ", cache_dir))
log_message(paste0("Use cache: ", use_cache))

# Initialize storage for PS results
twang_results <- list()

@@

for (i in 1:n_imputations) {
  # Check for cached result
  cache_file <- file.path(cache_dir, paste0("twang_imp_", i, ".rds"))

  if (use_cache && file.exists(cache_file)) {
    log_message(paste0("Imputation ", i, ": Loading from cache"))
    cached_result <- readRDS(cache_file)
    twang_results[[i]] <- cached_result
    weights <- cached_result$weights
  } else {
    log_message(paste0("Imputation ", i, ": Fitting twang model"))
    imp_start <- Sys.time()

    # Run twang ps model - following analysis.Rmd approach
    twang_model <- ps(
      formula = ps_formula,
      data = data.frame(imputed_datasets[[i]]),  # Convert to data.frame as in analysis.Rmd
      n.trees = n_trees,
      interaction.depth = 3,
      shrinkage = 0.01,
      estimand = "ATT",
      stop.method = c("es.mean"),  # Using es.mean only for speed
      verbose = FALSE
    )

@@

    # Store results
    result <- list(
      model = twang_model,
      weights = weights,
      n_trees_used = twang_model$desc$es.mean$n.trees
    )
    twang_results[[i]] <- result

    # Save to cache
    saveRDS(result, cache_file)

    imp_time <- round(as.numeric(difftime(Sys.time(), imp_start, units = "mins")), 2)
    log_message(paste0("Imputation ", i, ": Completed in ", imp_time, " minutes, saved to cache"))
  }

  # Track progress (works for both cached and newly computed)
  progress_df <- rbind(progress_df, data.frame(
    Imputation = i,
    Trees_Used = twang_results[[i]]$n_trees_used,
    Weight_Min = round(min(weights), 3),
    Weight_Max = round(max(weights), 3)
  ))

  # Log progress with time estimates
  log_progress(i, n_imputations, "Imputation")
}

@@

```{r save-results}
# Extract weights from all twang models for easy access in later steps
twang_weights <- list()
for (i in 1:n_imputations) {
  twang_weights[[i]] <- twang_results[[i]]$weights
}

@@

**Propensity score results saved to:** `r file.path(reanalysis_data_dir, "ps_results_twang.rds")`
**Weights saved to:** `r file.path(reanalysis_data_dir, "twang_weights.rds")`
**Cache directory:** `r cache_dir`

```{r cache-summary}
# Check cache status
cache_files <- list.files(cache_dir, pattern = "twang_imp_.*\\.rds", full.names = FALSE)
n_cached <- length(cache_files)

cache_status_msg <- paste0("Cache status: ", n_cached, " of ", n_imputations, " imputations cached")
log_message(cache_status_msg)

if (n_cached > 0) {
  log_message("To force re-computation, set use_cache <- FALSE in the setup chunk")
}

# Finalize logging
final_log <- finalize_log(success = TRUE)
```

**Log file location:** `r final_log`

To monitor the log during execution:
```bash
tail -f `r final_log`
```

### 2025-08-28 reanalysis-4-outcome_analysis.qmd
fix ps and implement quarto rstudio friendly logging

# Load data and weights
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
twang_weights <- readRDS(file.path(reanalysis_data_dir, "twang_weights.rds"))

@@

  # Initialize progress tracking
  progress <- init_progress(n_imputations, "Fitting models to imputed datasets")

@@

>>>>

# Initialize logging
log_file <- init_log("outcome_analysis")

# Load data and weights
log_message("Loading imputed datasets and weights")
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
twang_weights <- readRDS(file.path(reanalysis_data_dir, "twang_weights.rds"))

@@

  log_message(paste0("Starting model fitting for ", n_imputations, " imputations"))

  for (i in 1:n_imputations) {
    log_progress(i, n_imputations, "Model")

@@

```{r finalize-log}
# Finalize logging
log_message("Outcome analysis completed successfully")
final_log <- finalize_log(success = TRUE)
```

**Log file:** `r final_log`
